Andres Karjus

### 2020-11-05 index.Rmd
news

<div style="text-align: justify; padding-left:1px"> 
I am a PhD student in linguistics at the [Centre for Language Evolution at the University of Edinburgh](http://www.lel.ed.ac.uk/cle/), working on lexical dynamics and evolutionary variant selection in language with [Kenny Smith](http://www.lel.ed.ac.uk/~kenny), [Simon Kirby](http://www.lel.ed.ac.uk/~simon) and [Richard A. Blythe](http://www2.ph.ed.ac.uk/~rblythe3/), utilizing massive corpora and natural language processing to hunt traces of selection and competition in language; and experiments to see how individuals perform given limited linguistic resources. This research is supported by a Kristjan Jaak scholarship from the Archimedes Foundation of Estonia. I am also engaged in teaching stats and data visualization for the humanities and social sciences (see the tab above). On the right is the picture obligatory for all academic home pages to subtly suggest that in fact I do not spend _all_ my time in the office at all.

@@

*  Andres Karjus, Richard A. Blythe, Simon Kirby, Kenny Smith 2020. Communicative need modulates competition in language change | in review | [preprint](https://arxiv.org/abs/2006.09277)
* Andres Karjus, Richard A. Blythe, Simon Kirby, Kenny Smith 2020. Challenges in detecting evolutionary forces in language change using diachronic corpora. Glossa: a journal of general linguistics, 5(1), p.45. | [open access](http://doi.org/10.5334/gjgl.909) | [preprint](http://arxiv.org/abs/1811.01275) | [code](https://github.com/andreskarjus/wfsim_fit)
* Andres Karjus, Richard A. Blythe, Simon Kirby, Kenny Smith, 2020. Quantifying the dynamics of topical fluctuations in language. [Language Dynamics and Change](https://doi.org/10.1163/22105832-01001200) | [code](https://github.com/andreskarjus/topical_cultural_advection_model)
* Andres Karjus, Martin Ehala, 2018. Testing an agent based model of language choice on sociolinguistic survey data. Language Dynamics and Change, 8, pp. 219-252 | [Online at Brill](https://doi.org/10.1163/22105832-00802004) | [postprint](Karjus_Ehala_2018_Testing_an_agent-based_model_of_language_choice_LDCpostprint.pdf) | <a href="karjus_testing_2018.bib" target="_blank">bib</a> | sociolinguistic [dataset](http://doi.org/10.15155/1-00-0000-0000-0000-00162L): 1000 respondents, 200 questions
* Andres Karjus, Richard A. Blythe, Simon Kirby, Kenny Smith, 2018. Topical advection as a baseline model for diachronic lexical dynamics. Proceedings of The Society for Computation in Linguistics. Volume 1. [extended abstract, full paper above] | [open access](http://scholarworks.umass.edu/scil/) | <a href="karjus_topical_2018.bib" target="_blank">bib</a>
*	Martin Haspelmath, Andres Karjus, 2017. Explaining asymmetries in number marking: Singulatives, pluratives and usage frequency. Linguistics, volume 55, issue 6. | [online version](https://www.degruyter.com/view/j/ling.2017.55.issue-6/ling-2017-0026/ling-2017-0026.xml)  | [preprint](https://www.academia.edu/25032500/Explaining_asymmetries_in_number_marking_Singulatives_pluratives_and_usage_frequency) | <a href="haspelmath_explaining_2017.bib" target="_blank">bib</a>

>>>>

<div style="text-align: justify; padding-left:1px"> 
I am a research fellow at the [CUDAN](http://cudan.tlu.ee/) lab at Tallinn University, studying language and cultural dynamics. I recently defended my PhD student in linguistics at the [Centre for Language Evolution at the University of Edinburgh](http://www.lel.ed.ac.uk/cle/) on lexical competition and communicative needs in language, supervised by [Kenny Smith](http://www.lel.ed.ac.uk/~kenny), [Simon Kirby](http://www.lel.ed.ac.uk/~simon) and [Richard A. Blythe](http://www2.ph.ed.ac.uk/~rblythe3/).
I am also engaged in teaching stats and data visualization for the humanities and social sciences (see the tab above). On the right is the picture obligatory for all academic home pages to subtly suggest that in fact I do not spend _all_ my time in the office at all.

@@

* Competition, selection and communicative need in language change. PhD [thesis](https://andreskarjus.github.io/thesis/Andres_Karjus_phd_thesis.pdf), University of Edinburgh
*  Andres Karjus, Richard A. Blythe, Simon Kirby, Kenny Smith 2020. Communicative need modulates competition in language change | in review | [preprint](https://arxiv.org/abs/2006.09277)
* Andres Karjus, Richard A. Blythe, Simon Kirby, Kenny Smith 2020. Challenges in detecting evolutionary forces in language change using diachronic corpora. Glossa: a journal of general linguistics, 5(1), p.45. | [open access](http://doi.org/10.5334/gjgl.909) | [preprint](http://arxiv.org/abs/1811.01275) | [code](https://github.com/andreskarjus/wfsim_fit)
* Andres Karjus, Richard A. Blythe, Simon Kirby, Kenny Smith, 2020. Quantifying the dynamics of topical fluctuations in language. [Language Dynamics and Change](https://doi.org/10.1163/22105832-01001200) | [code](https://github.com/andreskarjus/topical_cultural_advection_model)
* Andres Karjus, Martin Ehala, 2018. Testing an agent based model of language choice on sociolinguistic survey data. Language Dynamics and Change, 8, pp. 219-252 | [Online at Brill](https://doi.org/10.1163/22105832-00802004) | [postprint](Karjus_Ehala_2018_Testing_an_agent-based_model_of_language_choice_LDCpostprint.pdf) | <a href="karjus_testing_2018.bib" target="_blank">bib</a> | sociolinguistic [dataset](http://doi.org/10.15155/1-00-0000-0000-0000-00162L): 1000 respondents, 200 questions
* Andres Karjus, Richard A. Blythe, Simon Kirby, Kenny Smith, 2018. Topical advection as a baseline model for diachronic lexical dynamics. Proceedings of The Society for Computation in Linguistics. Volume 1. [extended abstract, full paper above] | [open access](http://scholarworks.umass.edu/scil/) | <a href="karjus_topical_2018.bib" target="_blank">bib</a>
*	Martin Haspelmath, Andres Karjus, 2017. Explaining asymmetries in number marking: Singulatives, pluratives and usage frequency. Linguistics, volume 55, issue 6. | [online version](https://www.degruyter.com/view/j/ling.2017.55.issue-6/ling-2017-0026/ling-2017-0026.xml)  | [preprint](https://www.academia.edu/25032500/Explaining_asymmetries_in_number_marking_Singulatives_pluratives_and_usage_frequency) | <a href="haspelmath_explaining_2017.bib" target="_blank">bib</a>

### 2020-11-05 news.Rmd
news

@@

>>>>

- Defended my PhD [thesis](https://andreskarjus.github.io/thesis/Andres_Karjus_phd_thesis.pdf), "Competition, selection and communicative need in language change" and passed with minor corrections (29.10.2020)

- Started a postdoc at [CUDAN](http://cudan.tlu.ee/) lab at Tallinn University, provisionally as a visiting fellow until my defense (01.10.2020)

@@

- Delivered another dataviz workshop for the University of Edinburgh Centre for Data, Culture and Society (27.07.2020)
